\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-17 paper.qmd
Regenerate PBMC3K UMAP results by following the same steps as the author in Literature

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).

@@

@@

## Import data
df_2 <- read_rds("data/pbmc/pbmc.rds")
training_data_1 <- read_rds("data/pbmc/pbmc_training.rds")
test_1 <- read_rds("data/pbmc/pbmc_test.rds")

tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_umap.rds")
PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

@@

plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

plot_list2_pbmc <- plot_UMAP_2D(UMAP_pbmc) + #ggtitle("(b)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis5PBMC
#| out-width: 100%

plot_list1_pbmc + plot_list2_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
  plot_layout(ncol=5)
```

@@

aic_plot_pbmc + mse_plot_pbmc +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

>>>>

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
#| out-width: 100%

@@

### tSNE

# num_bins_tsne_gau <- calculate_effective_x_bins(.data = tSNE_data_gau, x = tSNE1,
#                                                 cell_area = 1) #56
num_bins_tsne_gau <- 16

shape_val_tsne_gau <- calculate_effective_shape_value(.data = tSNE_data_gau,
                                                      x = tSNE1, y = tSNE2)
## To extract bin centroids
hexbin_data_object_tsne_gau <- extract_hexbin_centroids(nldr_df = tSNE_data_gau, num_bins = num_bins_tsne_gau, shape_val = shape_val_tsne_gau, x = tSNE1, y = tSNE2)

df_bin_centroids_tsne_gau <- hexbin_data_object_tsne_gau$hexdf_data

tSNE_data_with_hb_id_gau <- tSNE_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_tsne_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_tsne_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), tSNE_data_with_hb_id_gau)

## Averaged on high-D
df_bin_tsne_gau <- avg_highD_data(.data = df_all_tsne_gau)

## Triangulate bin centroids
tr1_object_tsne_gau <- triangulate_bin_centroids(df_bin_centroids_tsne_gau, x, y)
tr_from_to_df_tsne_gau <- generate_edge_info(triangular_object = tr1_object_tsne_gau)

## Compute 2D distances
distance_tsne_gau <- cal_2D_dist(.data = tr_from_to_df_tsne_gau)

## To find the benchmark value
benchmark_tsne_gau <- find_benchmark_value(.data = distance_tsne_gau, distance_col = distance)

trimesh_removed_tsne_gau <- remove_long_edges(.data = distance_tsne_gau, benchmark_value = benchmark_tsne_gau,
                                                  triangular_object = tr1_object_tsne_gau, distance_col = distance)

trimesh_removed_tsne_gau <- trimesh_removed_tsne_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_tsne_gau <- show_langevitour(df_all_tsne_gau, df_bin_tsne_gau, df_bin_centroids_tsne_gau, benchmark_value = benchmark_tsne_gau, distance = distance_tsne_gau, distance_col = distance)

### UMAP

# num_bins_umap_gau <- calculate_effective_x_bins(.data = UMAP_data_gau, x = UMAP1,
#                                                 cell_area = 1) #41
num_bins_umap_gau <- 26

shape_val_umap_gau <- calculate_effective_shape_value(.data = UMAP_data_gau,
                                                      x = UMAP1, y = UMAP2)
## To extract bin centroids
hexbin_data_object_umap_gau <- extract_hexbin_centroids(nldr_df = UMAP_data_gau, num_bins = num_bins_umap_gau, shape_val = shape_val_umap_gau, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_gau <- hexbin_data_object_umap_gau$hexdf_data

UMAP_data_with_hb_id_gau <- UMAP_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_umap_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_umap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), UMAP_data_with_hb_id_gau)

## Averaged on high-D
df_bin_umap_gau <- avg_highD_data(.data = df_all_umap_gau)

## Triangulate bin centroids
tr1_object_umap_gau <- triangulate_bin_centroids(df_bin_centroids_umap_gau, x, y)
tr_from_to_df_umap_gau <- generate_edge_info(triangular_object = tr1_object_umap_gau)

## Compute 2D distances
distance_umap_gau <- cal_2D_dist(.data = tr_from_to_df_umap_gau)

## To find the benchmark value
benchmark_umap_gau <- find_benchmark_value(.data = distance_umap_gau, distance_col = distance)

trimesh_removed_umap_gau <- remove_long_edges(.data = distance_umap_gau, benchmark_value = benchmark_umap_gau,
                                              triangular_object = tr1_object_umap_gau, distance_col = distance)

trimesh_removed_umap_gau <- trimesh_removed_umap_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_umap_gau <- show_langevitour(df_all_umap_gau, df_bin_umap_gau, df_bin_centroids_umap_gau, benchmark_value = benchmark_umap_gau, distance = distance_umap_gau, distance_col = distance)

## PAHTE

# num_bins_phate_gau <- calculate_effective_x_bins(.data = PHATE_data_gau, x = PHATE1,
#                                                  cell_area = 1)
num_bins_phate_gau <- 70

shape_val_phate_gau <- calculate_effective_shape_value(.data = PHATE_data_gau,
                                                       x = PHATE1, y = PHATE2)
## To extract bin centroids
hexbin_data_object_phate_gau <- extract_hexbin_centroids(nldr_df = PHATE_data_gau, num_bins = num_bins_phate_gau, shape_val = shape_val_phate_gau, x = PHATE1, y = PHATE2)

df_bin_centroids_phate_gau <- hexbin_data_object_phate_gau$hexdf_data

PHATE_data_with_hb_id_gau <- PHATE_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_phate_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_phate_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PHATE_data_with_hb_id_gau)

## Averaged on high-D
df_bin_phate_gau <- avg_highD_data(.data = df_all_phate_gau)

## Triangulate bin centroids
tr1_object_phate_gau <- triangulate_bin_centroids(df_bin_centroids_phate_gau, x, y)
tr_from_to_df_phate_gau <- generate_edge_info(triangular_object = tr1_object_phate_gau)

## Compute 2D distances
distance_phate_gau <- cal_2D_dist(.data = tr_from_to_df_phate_gau)

## To find the benchmark value
benchmark_phate_gau <- find_benchmark_value(.data = distance_phate_gau, distance_col = distance)
benchmark_phate_gau <- 0.00078

trimesh_removed_phate_gau <- remove_long_edges(.data = distance_phate_gau, benchmark_value = benchmark_phate_gau,
                                               triangular_object = tr1_object_phate_gau, distance_col = distance)

trimesh_removed_phate_gau <- trimesh_removed_phate_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_phate_gau <- show_langevitour(df_all_phate_gau, df_bin_phate_gau, df_bin_centroids_phate_gau, benchmark_value = benchmark_phate_gau, distance = distance_phate_gau, distance_col = distance)

# num_bins_trimap_gau <- calculate_effective_x_bins(.data = TriMAP_data_gau, x = TriMAP1,
#                                                   cell_area = 1) #143

num_bins_trimap_gau <- 21
shape_val_trimap_gau <- calculate_effective_shape_value(.data = TriMAP_data_gau,
                                                        x = TriMAP1, y = TriMAP2)
## To extract bin centroids
hexbin_data_object_trimap_gau <- extract_hexbin_centroids(nldr_df = TriMAP_data_gau, num_bins = num_bins_trimap_gau, shape_val = shape_val_trimap_gau, x = TriMAP1, y = TriMAP2)

df_bin_centroids_trimap_gau <- hexbin_data_object_trimap_gau$hexdf_data

TriMAP_data_with_hb_id_gau <- TriMAP_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_trimap_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_trimap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), TriMAP_data_with_hb_id_gau)

## Averaged on high-D
df_bin_trimap_gau <- avg_highD_data(.data = df_all_trimap_gau)

## Triangulate bin centroids
tr1_object_trimap_gau <- triangulate_bin_centroids(df_bin_centroids_trimap_gau, x, y)
tr_from_to_df_trimap_gau <- generate_edge_info(triangular_object = tr1_object_trimap_gau)

## Compute 2D distances
distance_trimap_gau <- cal_2D_dist(.data = tr_from_to_df_trimap_gau)

## To find the benchmark value
benchmark_trimap_gau <- find_benchmark_value(.data = distance_trimap_gau, distance_col = distance)

trimesh_removed_trimap_gau <- remove_long_edges(.data = distance_trimap_gau, benchmark_value = benchmark_trimap_gau,
                                                triangular_object = tr1_object_trimap_gau, distance_col = distance)

trimesh_removed_trimap_gau <- trimesh_removed_trimap_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_trimap_gau <- show_langevitour(df_all_trimap_gau, df_bin_trimap_gau, df_bin_centroids_trimap_gau, benchmark_value = benchmark_trimap_gau, distance = distance_trimap_gau, distance_col = distance)

# num_bins_pacmap_gau <- calculate_effective_x_bins(.data = PaCMAP_data_gau, x = PaCMAP1,
#                                                   cell_area = 1) #30
num_bins_pacmap_gau <- 18

shape_val_pacmap_gau <- calculate_effective_shape_value(.data = PaCMAP_data_gau,
                                                        x = PaCMAP1, y = PaCMAP2)
## To extract bin centroids
hexbin_data_object_pacmap_gau <- extract_hexbin_centroids(nldr_df = PaCMAP_data_gau, num_bins = num_bins_pacmap_gau, shape_val = shape_val_pacmap_gau, x = PaCMAP1, y = PaCMAP2)

df_bin_centroids_pacmap_gau <- hexbin_data_object_pacmap_gau$hexdf_data

PaCMAP_data_with_hb_id_gau <- PaCMAP_data_gau |>
  dplyr::mutate(hb_id = hexbin_data_object_pacmap_gau$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_pacmap_gau <- dplyr::bind_cols(training_data_1 |> dplyr::select(-ID), PaCMAP_data_with_hb_id_gau)

## Averaged on high-D
df_bin_pacmap_gau <- avg_highD_data(.data = df_all_pacmap_gau)

## Triangulate bin centroids
tr1_object_pacmap_gau <- triangulate_bin_centroids(df_bin_centroids_pacmap_gau, x, y)
tr_from_to_df_pacmap_gau <- generate_edge_info(triangular_object = tr1_object_pacmap_gau)

## Compute 2D distances
distance_pacmap_gau <- cal_2D_dist(.data = tr_from_to_df_pacmap_gau)

## To find the benchmark value
benchmark_pacmap_gau <- find_benchmark_value(.data = distance_pacmap_gau, distance_col = distance)

trimesh_removed_pacmap_gau <- remove_long_edges(.data = distance_pacmap_gau, benchmark_value = benchmark_pacmap_gau,
                                                triangular_object = tr1_object_pacmap_gau, distance_col = distance)

trimesh_removed_pacmap_gau <- trimesh_removed_pacmap_gau +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.05) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_pacmap_gau <- show_langevitour(df_all_pacmap_gau, df_bin_pacmap_gau, df_bin_centroids_pacmap_gau, benchmark_value = benchmark_pacmap_gau, distance = distance_pacmap_gau, distance_col = distance)

```{r}
#| warning: false
#| echo: false
#| label: fig-modelfiveGau
#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).
#| out-width: 100%

trimesh_removed_tsne_gau + trimesh_removed_umap_gau + trimesh_removed_phate_gau + trimesh_removed_trimap_gau + trimesh_removed_pacmap_gau +
  #plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 5) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

@@

## Import data
#df_2 <- read_rds("data/pbmc/pbmc.rds")
training_data_1 <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_pca_50.rds")
training_data_1 <- training_data_1[, 1:15]
#test_1 <- read_rds("data/pbmc/pbmc_test.rds")

#tSNE_pbmc <- read_rds("data/pbmc/pbmc_tsne_30.rds")
UMAP_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_umap.rds")
# PHATE_pbmc <- read_rds("data/pbmc/pbmc_phate.rds")
# TriMAP_pbmc <- read_rds("data/pbmc/pbmc_trimap.rds")
# PaCMAP_pbmc <- read_rds("data/pbmc/pbmc_pacmap.rds")

load("data/pbmc/pbmc_3k_festem/pbmc3k_label.RData")

@@

# plot_list1_pbmc <- plot_tSNE_2D(tSNE_pbmc) + #ggtitle("(a)") + 
#   theme_linedraw() +
#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
#   annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

class_avg <- UMAP_pbmc |>
  group_by(cell_label) |>
  summarise(
    UMAP1 = median(UMAP1),
    UMAP2 = median(UMAP2)
  )

class_avg$cell_label <- as.factor(class_avg$cell_label)

levels(class_avg$cell_label) <- c("Memory \nCD4 T", "Naive CD4 T", "CD14+ Mono",  "B", "CD8 T", "FCGR3A+ \n Mono", "NK", "M-MDSC\n-like", "CD27-CD+ \n Memory T", "DC")

plot_list2_pbmc <- UMAP_pbmc |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = cell_label))+
    geom_point(alpha=0.5) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(legend.position = "none", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#b15928", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#6a3d9a", "#ff7f00", "#cab2d6", "#fdbf6f", "#ffff99", "#a6cee3", "#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#000000", "#bdbdbd")) +
  geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) #+
  # guides(color = guide_legend(title = "Cell type"),
  #        color = guide_legend(nrow = 3)) #+
  # annotate(geom = 'text', label = 'b', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# plot_list3_pbmc <- plot_PHATE_2D(PHATE_pbmc) + #ggtitle("(c)") + 
#   theme_linedraw() +
#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
#   annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# plot_list4_pbmc <- plot_TriMAP_2D(TriMAP_pbmc) + #ggtitle("(d)") + 
#   theme_linedraw() +
#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
#   annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)
# 
# plot_list5_pbmc <- plot_PaCMAP_2D(PaCMAP_pbmc) + #ggtitle("(e)") + 
#   theme_linedraw() +
#   theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
#         axis.title.x = element_blank(), axis.title.y = element_blank(),
#         axis.text.x = element_blank(), axis.ticks.x = element_blank(),
#         axis.text.y = element_blank(), axis.ticks.y = element_blank(),
#         panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
#   annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

```{r}
#| echo: false
#| fig-cap: "2D layout from UMAP (n_neighbors = 30, min_dist = 0.3) applied for the PBMC3k dataset. Is this a best representation of the original data?"
#| label: fig-nldervisPBMCUMAP

plot_list2_pbmc

#  plot_list1_pbmc + plot_list3_pbmc + plot_list4_pbmc + plot_list5_pbmc +
#   plot_layout(ncol=5) #| out-width: 100%

# 2D layouts from different NLDR techniques applied for the training PBMC dataset: (a) tSNE (perplexity = 30), (b) UMAP (n_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?
```

<!--
```{r}
#| warning: false
#| echo: false

@@

aic_plot_pbmc + mse_plot_pbmc +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```
-->

### 2024-01-17 paper.tex
Regenerate PBMC3K UMAP results by following the same steps as the author in Literature

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}

@@

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervis5PBMC-1.pdf}

@@

\caption{\label{fig-nldervis5PBMC}2D layouts from different NLDR
techniques applied for the training PBMC dataset: (a) tSNE (perplexity =
30), (b) UMAP (n\_neighbors = 15), (c) PHATE (knn = 5), (d) TriMAP
(n\_inliers = 5, n\_outliers = 4, n\_random = 3), and (e) PaCMAP
(n\_neighbors = 10, init = random, MN\_ratio = 0.9, FP\_ratio = 2). Is
there a best representation of the original data or are they all
providing equivalent information?}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltPBMC-1.pdf}

\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
different NLDR techniques applied to training PBMC dataset. What is the
best NLDR technique to represent the original data in 2D?}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelScurve-1.pdf}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-modelfiveGau-1.pdf}

\caption{\label{fig-modelfiveGau}Is there a best model to represent the
original data in 2D space or are they all providing equivalent
information?, (a) Model in the 2D space with tSNE
(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-nldervisPBMCUMAP-1.pdf}

@@

\caption{\label{fig-nldervisPBMCUMAP}2D layout from UMAP (n\_neighbors =
30, min\_dist = 0.3) applied for the PBMC3k dataset. Is this a best
representation of the original data?}
